Geospatial - Assignment 1 - Moritz Peist, Noemi Lucchi, Simon Vellin

### 2025-01-27 hw1/submission_notebook.Rmd
Added correct 2010 boundaries for paper 3

```{r}
library(osmdata)
library(sf)
library(spData)
library(tidyverse)
library(geodata)
library(readxl)
library(ggplot2)
library(raster)
library(terra)
```

@@

## Paper 3

@@

```{r}
# Read the .xls file with the Brasilian population for 2009  
df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)

# Get the geographical data from the package geobr
state <- read_state(year=2010)
meso <- read_meso_region(year=2010)
```

@@

# Create the column pop_share
df_merged <- df_merged %>%
  mutate(pop_share = Total_Estimada / sum(Total_Estimada)*100)

# Filter for North and Central-West regions
regions_to_highlight <- df_merged %>%
  filter(name_region %in% c("Centro Oeste", "Norte"))

# Join the geometries of these regions in a polygon 
regions_union <- regions_to_highlight %>%
  st_union() %>% 
  st_as_sf()
```

@@

```{r}
ggplot(data = df_merged) +
  geom_sf(aes(fill = pop_share), color = "white") +
  scale_fill_gradientn(
    colors = c("lightblue", "blue", "darkblue"),
    name = "Pop. Share",
    limits = c(0, 25)) +
  geom_sf(data = regions_union, fill = NA, color = "red", linewidth = 1)
```

In comparison the third paper map looked as follows: ![Figure x]()

>>>>

```{r}
library(osmdata)
library(sf)
library(spData)
library(tidyverse)
library(geodata)
library(readxl)
library(ggplot2)
library(raster)
library(terra)
library(lwgeom)
library(geobr)
```

@@

## [Paper 3](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3776990)

@@

```{r}
# Read the .xls file with the Brasilian population for 2009
df <- read_excel("data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)

# Get the geographical data from the package geobr
state <- read_state(year = 2010)
meso <- read_meso_region(year = 2010)
region <- read_region(year = 2010)
```

@@

# Create the column pop_share
df_merged <- df_merged %>%
  mutate(pop_share = Total_Estimada / sum(Total_Estimada) * 100)

# Aggregate regions as the author's did to get the correct boundaries
region <- region %>%
  mutate(
    name_region = case_when(
      name_region %in% c("Centro Oeste", "Norte") ~ 2,
      TRUE ~ 1
    )
  ) %>%
  dplyr::select(name_region, geom) %>%
  group_by(name_region) %>%
  summarise(geom = st_union(geom)) %>%
  sfheaders::sf_remove_holes()
```

@@

```{r}
# Final plot
ggplot(data = df_merged) +
  ggtitle("(c) 2010") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_sf(aes(fill = pop_share), color = "white") +
  scale_fill_gradientn(
    colors = c("lightblue", "blue", "darkblue"),
    name = "Pop. Share",
    limits = c(0, 25)
  ) +
  geom_sf(data = region, aes(color = factor(name_region)), fill = NA, linewidth = 1) +
  scale_color_manual(values = c("black", "red")) +
  guides(color = "none")
```

In comparison the third paper map looked as follows: ![Figure 4](data/paper3/original_plot.png)

Again, as outlined above, we only replicate the 2010 map, due to data non-availability in the prior periods.
